---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
________________________________________  
 ```{R Load Library}
 library(ggplot2)
 ```
  

### *1. Loading and preprocessing the data*
__________________________________________

The following section will load data.
```{R LoadData}
unzip(zipfile = "activity.zip")
data <- read.csv("activity.csv")
```


### *2. What is mean total number of steps taken per day?*
__________________________________________
```{R Calualation Part 1}
total_steps_day <- tapply(data$steps,data$date, sum, na.rm = TRUE)
qplot(total_steps_day, binwidth=1000,xlab = "total number of steps taken each day", ylab = "Frequency")
data_mean <- mean(total_steps_day, na.rm = TRUE)
data_median <- median(total_steps_day, na.rm = TRUE)
```

The mean of steps taken each day is : `r data_mean`   
The median steps taken for each day is : `r data_median`

### *3. What is the average daily activity pattern?*
__________________________________________
```{R Calulation Part 2}
average <- aggregate(list(steps = data$steps), by = list(interval = data$interval), FUN=mean, na.rm=TRUE)

ggplot(average, aes(interval, steps)) + geom_line(color = "blue", size = 0.7) + labs(title = "Average Daily Activity Chart", x = "5-minute intervals", y = "Average Number of Steps Taken")

max <- average[which.max(average$steps),]
print(max)
```

The interval that has the maximum number of steps is: `r max$interval`   


### *4. Imputing missing values*
__________________________________________
```{R Calculation Part 3}
result4a <- sum(is.na(data$steps))


```

Total number of missing values is `r result4a`.

```{R Devise a inserting strategy}
insert_func <- function(interval,steps){
        if(is.na(steps)){
              average[which(average$interval==interval),"steps"]
        }else{
              steps
        }
}
```


```{R inserting value}
filled_data <- data
filled_data$steps  <- mapply(insert_func,filled_data$interval, filled_data$steps)

```

Create Histogram

```{r create histogram}
totalSteps <- tapply(filled_data$steps, filled_data$date, FUN=sum)
qplot(totalSteps, binwidth = 1000, xlab = "total number of steps taken each day")

```

```{r}

na_mean <- mean(totalSteps, na.rm = TRUE)
na_mean
na_median <- median(totalSteps, na.rm=TRUE)
na_median

```


### *5. Are there differences in activity patterns between weekdays and weekends?*
__________________________________________





